---
title: "Week 2 Exercises"
author: "Amanda Frithsen"
date: "July 8, 2024"
output: pdf_document
---

Please complete all exercises below. You may use stringr, lubridate, or the forcats library. 

Place this at the top of your script:
library(stringr)
library(lubridate)
library(forcats)

# Exercise 1
Read the sales_pipe.txt file into an R data frame as sales. 
```{r}
library(stringr)
library(lubridate)
library(forcats)
```


```{r}
sales <- read.delim("sales_pipe.txt"
                     ,stringsAsFactors=FALSE
                     ,sep = "|"
                     ,fileEncoding = "UTF-8"
                    )

```

# Exercise 2
You can extract a vector of columns names from a data frame using the colnames() function. Notice the first column has some odd characters. Change the column name for the FIRST column in the sales date frame to Row.ID. 

**Note: You will need to assign the first element of colnames to a single character.**

```{r}
colnames(sales)

colnames(sales)[1] <- "Row.ID"

colnames(sales)
```

# Exercise 3
Convert both Ship.Date and Order.Date to date vectors within the sales data frame. What is the number of days between the most recent order and the oldest order?
How many years is that?
How many weeks?

**Note: Use lubridate**

```{r}

# Convert Ship.Date to date vector
inherits(sales$Ship.Date, c("Date"))
sales$Ship.Date <- mdy(sales$Ship.Date)
inherits(sales$Ship.Date, c("Date"))

# Convert Order.Date to date vector
inherits(sales$Order.Date, c("Date"))
sales$Order.Date <- as.Date(sales$Order.Date, , "%m/%d/%Y")
inherits(sales$Order.Date, c("Date"))

# Order date of most recent order and oldest order
newest_order <- sales$Order.Date[which.max(sales$Order.Date)]
oldest_order <- sales$Order.Date[which.min(sales$Order.Date)]

# There are 1,457 days between the most recent order and the oldest order.
# There are 3 years, 11 months, 27 days between the most recent order and the oldest order.
# There are 208 weeks, 1 day between the most recent order and the oldest order. 

order_span <- interval(ymd(oldest_order), ymd(newest_order))
order_span

days <- as.period(order_span, unit = "days")
days

years <- as.period(order_span, unit = "years")
years

weeks <- interval(ymd(oldest_order), ymd(newest_order))/weeks(1)
weeks
```


# Exercise 4
What is the average number of days it takes to ship an order?

```{r}
diff_time <- interval(ymd(sales$Order.Date), ymd(sales$Ship.Date))/days(1)
average_ship_days <- sum(diff_time)/length(diff_time)
average_ship_days

# The average number of days it takes to ship an order is approximately 3.91 days.
```


# Exercise 5
How many customers have the first name Bill?
You will need to split the customer name into first and last name segments and then use a regular expression to match the first name bill. Use the length() function to determine the number of customers with the first name Bill in the sales data. 

```{r}
temp_name <- str_split_fixed(string = sales$Customer.Name,
                             pattern = " ", n = 2)
sales$Customer.First.Name <- temp_name[,1]
sales$Customer.Last.Name <- temp_name[,2]
head(sales)

sales <- sales[, c(1:6, 23, 22, 8:21)]
head(sales)

named_bill <- sales[sales$Customer.First.Name == "Bill", ]
length(named_bill)
length(unique(named_bill$Customer.Last.Name))

# There are 22 customers with the first name Bill.  
# There are 6 unique customers with the first name Bill.
```

# Exercise 6
How many mentions of the word 'table' are there in the Product.Name column? 
**Note you can do this in one line of code**

```{r}
mentions_of_table <- length(str_subset(sales$Product.Name, regex(" table", ignore_case = TRUE)))
mentions_of_table

# I included the method below as I was not fully trusting my first method and wanted to double.  I did need to run two lines of code in order to ensure that all characters were lower case.

sales$Product.Name <- str_to_lower(sales$Product.Name)
mentions_of_table_confirmed <- sum(str_detect(sales$Product.Name, " table"))
mentions_of_table_confirmed

# There are 246 mentions of the word 'table' in the Product.Name column.
```

# Exercise 7
Create a table of counts for each state in the sales data. The counts table should be ordered alphabetically from A to Z. 

```{r}
sales$State <- factor(sales$State)
is.factor(sales$State)
levels(sales$State)

state_table <- table(sales$State)
state_table
```

# Exercise 8
Create an alphabetically ordered barplot for each sales Category in the State of Texas. 

```{r}
sales$Category <- factor(sales$Category)
is.factor(sales$Category)
levels(sales$Category)

Texas <- sales[sales$State == "Texas",]
barplot(table(Texas$Category))
```

# Exercise 9
Find the average profit by region.
**Note: You will need to use the aggregate() function to do this. To understand how the function works type ?aggregate in the console.**

```{r}

average_profit_by_region <- aggregate(sales$Profit, list(Region = sales$Region), mean)
average_profit_by_region
```

# Exercise 10
Find the average profit by order year. 
**Note: You will need to use the aggregate() function to do this. To understand how the function works type ?aggregate in the console.**

```{r}
sales$Order.Year <- year(sales$Order.Date)

average_profit_by_year <- aggregate(sales$Profit, list(Year = sales$Order.Year), mean)
average_profit_by_year
```

